---
title: "ReoData_PeerAssessment1"
author: "Mark Oost"
date: "Sunday, August 17, 2014"
output: html_document
---

# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
1. Load the data.hjjhhjhhj
2. Process/transform the data into a format suitable for your analysis.


```{r, Loading Data}
#Loading data
activity <- read.csv("activity.csv")
#Convert variable to date/time
activity$date1 <- as.Date(activity$date, format = "%Y-%m-%d")
summary(activity)
```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day.


```{r, Create plot 1}
#Create plot data
plotdata <- tapply(activity$steps, activity$date, FUN = sum)
#Create plot
hist(plotdata, main = "Histogram", xlab = "number of steps taken each day", ylab ="number of times this total was reached")
```

2. Calculate and report the mean and median total number of steps taken per day.


```{r, Aggregrate data 1}
#Aggregrating data before calculating the mean
Aggregrate <- aggregate(steps ~ date, data = activity, FUN= sum, na.action = na.omit )
```

*mean and median*

```{r, Calculate Mean 1}
#Calculate mean and median
mean(Aggregrate$steps)
```

```{r, Calculate Median 1}
median(Aggregrate$steps)
```


## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.


```{r, Create plot 2}
#Creating data for plot
AggDailyPattern <- aggregate(steps ~ interval, data = activity, FUN= mean, na.action = na.omit )
#Plotting the data
plot(AggDailyPattern$interval, AggDailyPattern$steps, type = "l", xlab ="interval", ylab = "steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
*See answer below*


```{r, Sort data and take first}
#Sort data descending and take first
Sort <- AggDailyPattern[order(-AggDailyPattern$steps),]
Sort[1,]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset


```{r, Summarizing missings}
#SUmmarizing missings
sum(is.na(activity))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean median for that day, or the mean for that 5-minute interval, etc.
*I used the inverval avarage for imputing the variabels and have less unknowns and a better spread*

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r, Imputing Missings}
#Create new dataset
activitynomissing <- activity
#Imputing missings
activitynomissing$steps[is.na(activitynomissing$steps)] <- AggDailyPattern$steps
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment What is the impact of imputing missing data on the estimates of the total daily number of steps?
*I used the inverval avarage for imputing the variabels and have less unknowns and mean and median are now the same*


```{r, Create Plot 3}
#Create the data
plotdata <- tapply(activitynomissing$steps, activitynomissing$date, FUN = sum)
#Plot the data
hist(plotdata, main = "Histogram", xlab = "number of steps taken each day", ylab ="number of times this total was reached")
```

```{r, Aggregrate data 2}
#Aggregrating data before calculating the mean
Aggregratenomissing <- aggregate(steps ~ date, data = activitynomissing, FUN= sum, na.action = na.omit )
```


*mean and median after imputing*

```{r, Calculate Mean 2}
#Calculate mean and median
mean(Aggregratenomissing$steps)
```

```{r, Calculate Median 2}
median(Aggregratenomissing$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levelsweekday and weekend indicating whether a given date is a weekday or weekend day.


```{r, Create weekdays}
#Creating weekdays
activitynomissing$days <- weekdays(activitynomissing$date1)


#Create difference between weekday and weekend and make aggregrate
activitynomissing$weekend <- 
  ifelse(activitynomissing$days=='Sunday',"weekend",
         ifelse(activitynomissing$days=='Saturday',"weekend","weekday"))

weekend <- activitynomissing[activitynomissing$weekend=='weekend',]
weekendagg <- aggregate(steps ~ interval, data = weekend, FUN= mean)
weekendagg$label <- "weekend"
weekday <- activitynomissing[activitynomissing$weekend=='weekday',]
weekdayagg <- aggregate(steps ~ interval, data = weekday, FUN= mean)
weekdayagg$label <- "weekday"
```

2. Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days. The plot should look something like the following, which was creating using simulated data:


```{r, Create plot 4}
#Creating plotdata
plotdata <- rbind(weekendagg, weekdayagg)

#Loading lattice and creating plot
library(lattice)
xyplot(steps ~ interval | label, data=plotdata, type="l", layout=c(1,2))
```